---
title: "Commute Convergence Vignette"
author: "David Burton"
date: "December 14, 2017"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

```

# Basic Commute Analysis

So, you've been driving to the same place more and more often, but it's eating you
alive inside that you don't know when it will take you the least time to get there.

This R package is for you. _commuteconvergence_ is a simple package that has a very simple
requirement: you must record the day of the week and total commute time each day, or when you
leave home and arrive at work, in the columns of a Google Sheet.

_commuteconvergence_ will harness the power of the great _googlesheets_ and _ggplot2_ 
packages to work for you!

##Step 0: Collect data    

It is *extremely* important that you record your data into a Google Sheet as instructed.
At the bare minimum it must include two columns: Day and Total(drive time each day). 
Day is a group by which you want to compare your times, so you can use other groups than day of
the week, such as what you listen to in the  car. However, you also have the 
option to have the package calculate drive time for you if you will supply a Leave time 
and Arrive time instead of Total. Leave and Arrive must be formatted in HH:MM:SS; such as 
10:23:00 AM, for 10:23 AM. Googlesheets seems to do this automatically, but be careful to double
check if you report afternoon and morning times. Using actual times also provides you more in depth, 
and probably, more useful analysis.

##Step 1: Read in data   

Get _commuteconvergence_ going using *library()*.
After you publish your Google sheet to the web for viewing, via the file menu in Google Sheets,
use the share button to get a url. Paste it in the function *read_sheet()*.
If unsure of how to share via a link in Google Sheets please consult a Google help
resource, or view documentation related to the _googleheets_ package.
You have to place your own quotation marks around the sharing link.
Assign the value of *read_sheet* to the data frame name you have chosen.    

I have included a link to an example minimal sheet which is available for public viewing
on Google Sheets. Please replace it with your own data once you understand the process.   
```{r message=FALSE}
library(commuteconvergence)
commute <- read_sheet(
"https://docs.google.com/spreadsheets/d/14TN5tRLf2HQq3m8Fot8VhEydobYqDW_7hBWqLWPcQzw/edit?usp=sharing")

```

##Step 2: View the data   
Using the *commuteplots* function with your newly minted dataframe,
one line of code will net you two relatively, well formatted graphs of your untransformed 
commute data.   

The defaults assume that you want to look at your data comparing total drive time
by day of the week. Thus, the _Time_ variable = "Total", and the _Group_ variable ="Day".
```{r}
commuteplots(commute)
```     
   
If you don't want to use days of the week, just subsitute in for _Group_. 
For instance, here I have run the plots based on 
comparing my total drive times by what I was listening to that morning. 
I also used actual times instead of a total drive time. This will net you 
another plot of total drive time explained by what time you left.

```{r message=FALSE}
davscommute <- read_sheet(
  "https://docs.google.com/spreadsheets/d/12-8vjn46tIZ2Xr98_o_AcUzX5qeMAV878h63wFPo4d4/edit?usp=sharing")

commuteplots(davscommute,Leave = "Left",Arrive="ArriveDesk",Group="Listening")

```   

    
##Step 3: Find out if times are significantly different.   
Using the *commuteanova* function on your data frame will let you know if there
are any significant differences based on the day you are driving. If no actual times are used,
 assumptions of ANOVA are ignored, and an ANOVA is run based on day of the week. 

```{r}
commuteanova(commute)
```
   
###What about for days that are significantly different?   
For that we'll need a Tukey Post-Hoc test and some different input to demonstrate.   

Again, here is a practice link of data to use. It is artificically crafted to be
significant for demonstration purposes.   
```{r message=FALSE}
different <- read_sheet(
"https://docs.google.com/spreadsheets/d/1d_KjDBLt-iSiRcGDvWsGQMNUlrPselEwjCN60gwvfkM/edit?usp=sharing")

commuteplots(different)
commuteanova(different)
```    

Here is an example using real data with actual times.
This gives you a k-means clustering of the times you leave
as well as an ANOVA to tell you if those clusters are
significantly different from one another. _usertimebins_ controls
how many clusters are used.

For the sake of the vignette, it can not be run in a chunk, since the resulting error
will stop the markdown from knitting.

###commuteanova(davscommute,Leave="Left",Arrive="ArriveDesk",usertimebins=6)
   

[1] "Leave between 09:48:00 and 11:02:00 for the shortest average drive time based on descriptive st... <truncated>
 Show Traceback
Error in commuteanova(davscommute, Leave = "Left", Arrive = "ArriveDesk", : No formal statistical test was conducted as ANOVA assumptions were not met by the data.



Built in error messages, such as the one above, are in place to prevent an ANOVA being 
conducted when assumptions of an ANOVA
test are not met. These include normally distributed data and homogeneity of variance.

Therein lies the problem though. Commute data seems likely to include ties, and in this
case non-parameteric rank tests are unvailable due to ties. It seems then that descriptive statistics
may be the only way to truly analyze the data. In this case, the lowest mean and variance seem
to be desirable, with mean probably having greater significance.

##Step 4: Watch for future updates.   

Future patches may include different types of statistical tests built to handle the data type better. 
Then you'll be extra certain you know which 
times or days you should drive instead of working at home. Naturally, though, you'll have some 
type of mandatory meeting that puts you on the road for your least optimal drive time, but this 
package can't do much about that! ;)
    

##Recap

Leaving between 8:46 and 9:31 seemed to give me the best result. The mean was pretty low, and the variance
of the time group was small with one outlier at 35 minutes. Other groups did seem to perform better, but
there also needs to be some consideration given to when I needed to be at Saunders. Unless I wanted to leave
very early in the morning, or arrive very late, this seemed to be the best compromise. Unfortunately, 
Probability Theory was at 9:15, and this guaranteed that at least two days a week I was on the road 
during the worst times, 7:18 to 8:02. Although, these interpretations are highly subjective since
interpreting the boxplots is the only valid course of action at this point.

If I had it to do over again, I may have tried to synthesize some kind of test statistic based around
variance and mean. It's pretty obvious now that this data didn't lend itself to my chosen method of
analysis. If I were going to continue working on this, I would probably move in that direction. I may
have tried to come up with another type of data to examine. My goals were so ambitious that there are
a lot of logic checks built into the package that may have been unnecessary. They also took a lot of time.


```{r}
sessionInfo()
```
